Consider a quarterly model of US monetary policy. Let $y_t=(\Delta p_t, \Delta gnp_t, i_t, \Delta m_t)'$, where $p_t$ refers to the log of the GNP deflator, $gnp_t$ to toe log of real GNP, $i_t$ to the federal funds rate, averaged by quarter, and $m_t$ to the log of M1.
The data is given in the excel sheet \texttt{Keating1992} of \texttt{data.xlsx}. The structural shock vector $\varepsilon_t = (\varepsilon_t^{AS},\varepsilon_t^{IS},\varepsilon_t^{MS},\varepsilon_t^{MD})$ includes an aggregate supply shock, an IS shock, a money supply shock, and a money demand shock. The structural model can be written as:
$$\begin{pmatrix}
        u_t^p \\u_t^{gnp}\\u_t^{i}\\u_t^m
    \end{pmatrix} =
    \begin{pmatrix}
        \varepsilon_{t}^{AS}                                                 \\
        -b_{21,0}u_t^p - b_{23,0} u_t^i- b_{24,0}u_t^m +\varepsilon_{t}^{IS} \\
        -b_{34,0}u_t^m + \varepsilon_{t}^{MS}                                \\
        -b_{41,0}(u_t^{gdp}+u_t^p)-b_{43,0}u_t^i +\varepsilon_{t}^{MD}
    \end{pmatrix}$$
Furthermore, it is assumed that $E(\varepsilon_t \varepsilon_t')=\Sigma_\varepsilon$ is not the identity matrix and hence the diagonal elements of $B_0$ equal unity.
\begin{enumerate}
    \item Estimate a VAR model with four lags and an intercept.
    \item Provide intuition behind the structural model and derive the $B_0$ matrix. Note that we are imposing restrictions on both $B_0$ and $\Sigma_\varepsilon$ but not on $B_0^{-1}$.
    \item Estimate $B_0$ and $\Sigma_\varepsilon$ using the nonlinear equation solver fsolve, i.e. first set up a function that computes
          $$ F_{SR}([B_0, diag(\Sigma_\varepsilon)]) = \begin{bmatrix} vech\left(B_0^{-1} \Sigma_\varepsilon B_0^{-1'} - \hat{\Sigma}_u\right)\\\text{restrictions on }B_0\end{bmatrix}=0$$
          where $diag(\Sigma_\varepsilon)$ denotes only the diagonal elements of $\Sigma_\varepsilon$. Note that these diagonal elements as well as $B_0$ are the inputs stacked in a matrix or vector for the optimizer.
          Use the same options for the optimizer as in the previous exercise.
    \item Use the estimated structural impact matrix $B_0^{-1}\Sigma_\varepsilon^{1/2}$ to plot the structural impulse response functions using \texttt{IRFs.m}. Interpret the effects of an aggregate supply shock on prices, real GNP, Federal Funds Rate and M1.
\end{enumerate}
\newpage
